
---
title: "WAreg: A Regression Framework for Weighted Composite Endpoints"
author: "Xi Fang"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{WAreg Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Introduction

Composite endpoints are frequently used in clinical trials to enhance event incidence and improve statistical power.The **WAreg** package provides a regression framework for weighted composite endpoints using "while-alive" cumulative frequency measures. This approach incorporates time-varying effects using splines and is applicable to both conventional and cluster-randomized trials. In this vignette we demonstrate how to:

- Load an example dataset.
- Fit a WA regression model.
- Summarize the fitted model.
- Visualize the estimated time-varying effects.
- (Optionally) Perform cross-validation to select model complexity.

# Loading the Example Data

The package includes an example dataset, `ex_dt`, stored in the `data/` folder. Load it using:

```{r load-data}
data(ex_dt)
head(ex_dt)
```

The dataset contains columns such as `Cluster`, `Subject`, `time`, `type`, `delta`, `Z1`, and `Z2`.

# Fitting the WA Regression Model

The main function for fitting the model is `WA_reg()`, which integrates functions defined in `util_fun.R`, `WA_fit.R`, and `WA_fun.R`. For example, suppose we have the following model specifications:

```r
model <- Surv(time, type) ~ Z1 + Z2
cens_mod <- Surv(time, delta) ~ Z2
```

Other parameters include:
  
- `wr`: Weight vector for recurrent events.
- `wd`: Weight for the terminal event.
- `sp_knots`: Spline knot locations, e.g., `c(5,10,15,20,25,30,35,40)`.
- `degree`: Spline polynomial degree.
- `id`: Subject ID column name (e.g., `"Subject"`).
- `cluster`: Cluster column name (e.g., `"Cluster"`).
- `ts`: A vector of time points for evaluation (e.g., `seq(0.001,40,length.out = 1000)`).

Fit the model as follows:

```{r fit-model}
# Set parameters
wr <- c(1, 1)
wd <- 1
sp_knots <- c(5, 10, 15, 20, 25, 30, 35, 40)
degree <- 1
id <- "Subject"
cluster <- "Cluster"
ts <- seq(0.001, 40, length.out = 1000)

# Fit the WA regression model. WA_reg() is defined in the package.
fit_obj <- WA_reg(model = model, data = ex_dt, cluster = cluster, id = id,
                  wr = wr, wd = wd, cens_mod = cens_mod,
                  sp_knots = sp_knots, degree = degree,
                  cv = FALSE, K = 10, tr = c(5,35),
                  nk = 2:6, ts = ts)
```

# Summarizing the Model

```{r summary-model}
summary(fit_obj)
```

# Plotting the Time-Varying Effects

```{r plot-results}
plot_results <- plot.WA_obj(fit_obj, ts = seq(0, 40, by = 1), alpha = 0.05)
```

# Cross-Validation (Optional)

```{r cross-validation, eval=FALSE}
cv_error <- cv_fit(model = model, data = ex_dt, cluster = cluster, id = id,
                   wr = wr, wd = wd, cens_mod = cens_mod,
                   degree = degree, time_range = sp_knots,
                   L = ex_dt$N_R, K = 10, t_span = ts)
cat("Cross-validation error:", cv_error)
```

# Conclusion

This tutorial demonstrated the full workflow of the **WAreg** package: from loading example data, fitting a weighted composite endpoint regression model with time-varying effects using splines, to summarizing and visualizing the results. For more details on the methodology and further examples, please refer to the package documentation.
